Exploring Local Explanations of Non-linear Models Using Animated Linear Projections

### 2022-01-11 _paper.rmd
chocolates changed to vary fiber from sugar.

(ref:casechocolates) Chocolates data type classification (milk or dark). We select a chocolate labeled as dark though a random forest model predicts it to be milk chocolate from the values on the nutritional label. The attribution projection already looks more like a dark chocolate than milk. We remove a 4 variables with the lowest contribution for the selected instance and vary the contribution of `Sugar`. The misclassification of two instances seems improbable when sugar is near max contribution. Animated tour can be found at [shorturl.at/dBE28](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_chocolates.mp4).
```{r casechocolates, echo=F, out.width="100%", fig.cap = "(ref:casechocolates)"}
knitr::include_graphics("./figures/case_chocolates.png")
```

Figure \@ref(fig:casechocolates) similarly shows that attribution-space is more separable than data-space. Interestingly, the class imbalance that we suspected was not observed; there are only six chocolates labeled as dark and predicted as milk, while eight of the inverse case. We look at one extreme instance and vary the contribution of `sugar`. We see notice 2 instances that have clearly well within the milk chocolate dispite being observed as dark chocolate on their product label. While conjector, there are plausible reasons that a manufacturer has incentives to cut corners and label their products different than what they are.

@@

@@

(ref:caseames) Ames housing 2018 regressing sales price [2018 USD]. The primary observation is a house that a large `lot area` (`LtA`), but small `living area` (`LvA`). The comparison instance is in the opposite; a large house on a relatively small lot. Similar to the soccer players, these house sale have diverging attribution to better predict their sales prices. Varying the contribution `living area` the separation between these house sales crosses when there is a low contribution which is so important to explaining their prediction though in differet ways. The corresponding animation is at [shorturl.at/oqyDU](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_ames2018.mp4)
```{r caseames, echo=F, out.width="100%", fig.cap = "(ref:caseames)"}
knitr::include_graphics("./figures/case_ames2018.png")
```

>>>>

XXX: NS: not as bad as I remember, but a couple, remove them?
```{r}
cheem::chocolates$Name[c(26, 74)]
```

(ref:casechocolates) Chocolates data type classification (milk or dark). We select a chocolate labeled as dark though a random forest model predicts it to be milk chocolate from the values on the nutritional label. The attribution projection already looks more like a dark chocolate than milk. We remove a 4 variables with the lowest contribution for the selected instance and vary the contribution of `fiber`. The misclassification of two instances seems improbable when sugar is near max contribution. Animated tour can be found at [shorturl.at/dBE28](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_chocolates.mp4).
```{r casechocolates, echo=F, out.width="100%", fig.cap = "(ref:casechocolates)"}
knitr::include_graphics("./figures/case_chocolates.png")
```

Figure \@ref(fig:casechocolates) similarly shows that attribution-space is more separable than data-space. Interestingly, the class imbalance that we suspected was not observed; there are only six chocolates labeled as dark and predicted as milk, while eight of the inverse case. We look at one extreme instance and vary the contribution of `fiber`. We see notice two instances that have clearly well within the milk chocolate despite being observed as dark chocolate on their product label. While conjecture, there are plausible reasons that a manufacturer has incentives to cut corners and label their products different than what they are.

@@

NS: Sugar is slightly better imo, but fiber is fine.

@@

(ref:caseames) Ames housing 2018 regressing sales price [2018 USD]. The primary observation is a house that a large `lot area` (`LtA`), but small `living area` (`LvA`). The comparison instance is in the opposite; a large house on a relatively small lot. Similar to the soccer players, these house sale have diverging attribution to better predict their sales prices. Varying the contribution `living area` the separation between these house sales crosses when there is a low contribution which is so important to explaining their prediction though in different ways. The corresponding animation is at [shorturl.at/oqyDU](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_ames2018.mp4)
```{r caseames, echo=F, out.width="100%", fig.cap = "(ref:caseames)"}
knitr::include_graphics("./figures/case_ames2018.png")
```
